Learning from Data Through Models"  # nolint: Grammarly.

### 2023-09-07 Paper/draft.Rmd
editing text + fixing plots

# Setting 1

@@

$$\frac{p[h^t|\tilde{\theta}]}{p[h^t|\hat{\theta}]} > \alpha$$

Where $h^t$ is the history of outcomes up to time $t$ and $\alpha \geq 1$ is a threshold that determines how convincing the data needs to be
for the agent to change their belief. Notice that if $\alpha \to \infty$, we get the dogmatic agent. In this sense, the 
switcher is a generalization of the dogmatic type, just as the self-attribution agent is a generalization of the Bayesian agent.

By allowing the agent to keep track of the likelihoods and switching to an alternative type, the switcher can avoid the 
self-confirming equilibrium the dogmatic agent falls into. However, if the prior belief on $theta$ is sufficiently tight around
a self-confirming equilibrium, the switcher might look identical to the dogmatic even in a case where $\alpha$ is not too large.

** Example 1 (continued)** 
_ Take the same student from the example and now add the possibility of switching beliefs. _

# Seting 2

@@

Bayesian agents always choose the effort level that maximizes their flow payoff by taking expectations over
their prior beliefs about $\theta$ and $\omega$. Since agents are myopic, even though all the parameters could be identified with enough variation in choices,

a fully Bayesian agent might not learn their true type. This happens because they do not internalize the tradeoff between flow payoff
and learning and thus might not experiment enough to learn their type. An alternative to this approach is given 
by @Hestermann2021 and is discussed with the results.

@@

# A Unifying Framework
The agent can be of one of 3 types: $\theta \in \{\theta_L, \theta_M, \theta_H\}$  with $\theta_H > \theta_M > \theta_L$.
They face an unknown exogenous success rate
$\omega \in \{\omega_L, \omega_M, \omega_H\}$ with $\omega_H>\omega_M>\omega_L$. Each of the values of $\omega$ is realized with equal probability. 
The agent knows the distribution of $\omega$ but not its realized value. Let the true type be $\theta^*$ 
and let the true state be $\omega^*$.

The agent holds some prior belief about $\theta$ \footnote{which is potentially misspecified as in the dogmatic and switcher cases discussed} 
and chooses a binary gamble $e in \{e_L, e_M, e_H\}$. The agent observes whether the gamble is a success or a failure and 
gets a payoff of $1$ and if it is a success. They get $0$ otherwise.

The probability of success is increasing in
both $\theta$ and $\omega$ and is fully described in the following table:

@@

Conditional on a type, the agent's flow payoff is maximized by choosing the gamble that 
matches the state. For example, if the value of $\omega$ is $\omega_H$, the agent's flow payoff is maximized by 
choosing $e_H$ and if the state is $\omega_L$ the flow payoff is maximized by choosing gamble $e_L$. The agent myopically chooses
gambles every period to maximize the flow payoff for $T$ periods.

After observing the outcome of each gamble, the agent updates their beliefs using some algorithm and moves on to the next period.

Notice that both $\theta$ and $\omega$ can be identified from the outcomes if enough variation in the effort choices exists.
This can be seen by confirming there is no pair of $\theta$ and $\omega$ such that the probability of success is the same for all effort 
choices. Therefore, by changing the effort choice, the agent can learn both their type and the state if they observe enough outcomes.

In this example, for an agent with a dogmatic belief about their type, a self-defeating equilibrium is one in which the agent 
chooses an effort level that, under the true $\theta$, yields a frequency of success that is consistent with the agent's 
misspecified belief. That is, the observed rate of success under the true type and the effort level chosen, 
$P[\text{sucess}|\theta^*, e]$ is the same as the probability of success under the agent's belief, $P[\text{sucess}|\hat{\theta}, e]$.

In the data-generating process described above, there are five such equilibria. For example, if the agent is of type $\theta_M$  but
mistakenly believes that he is of type $\hat{\theta}=\theta_H$ and the 
and $\omega^* = \omega_M$, when the effort chosen is $e_L$, the agent will observe a success with 45% chance. Because the agent 
dogmatically believes that their type is high, they will erroneously conclude that the rate is $\omega_L$. Under this belief,
the optimal action is $e_L$ which will continue to generate successes with $45%$ probability, further reinforcing
the incorrect belief. If the agent were to learn the true type and rate, they would instead choose $e_M$, yielding a 
success with $65%$ probability and strictly higher expected payoffs.

From @Heidhues2018, we know that if $\theta^*$ and $\omega^*$ are such that there is a self-confirming equilibrium, a dogmatic agent will always
fall into the trap and continue to make sub-optimal choices. On the other hand, a switcher will be able to learn the true type and state
as long as their prior is diffused enough and their switching threshold is low enough as in @Ba2023.

By including self-confirming equilibria, the example captures the forces from each of the updating mechanisms discussed in the 
previous section and allows for direct comparison of all the theories. For realizations of $(\theta, \omega)$ for which there are 
self-confirming equilibria, the dogmatic agent will fall into the 
trap whereas the switcher will be able to escape it. Similarly, an agent with self-attribution bias will update their beliefs differently from an unbiased Bayesian, leading
them to choose different gambles. I exploit such cases in order to test which model is a better fit for how subjects behave in a 
laboratory experiment.

ILLUSTRATE THE DIFFERENT PATHS WITH THE SIMULATION.

A dogmatic modeler updates their beliefs about omega by using bayes rule. For the example, that means he ends up with the following
posterior odds ratios:

\begin{equation}
\frac{p_{t}[\omega_H|\text{outcome}]}{p_{t}[\omega_M|\text{outcome}]} = 
      \frac{p[\text{outcome}|\omega_H, \hat{\theta}, e]p_{t-1}[\omega_H]}{p[\text{outcome}|\omega_M, \hat{\theta}, e]p_{t-1}[\omega_M]}
\end{equation}
and
\begin{equation}
\frac{p_{t}[\omega_M|\text{outcome}]}{p_{t}[\omega_L|\text{outcome}]} = 
      \frac{p[\text{outcome}|\omega_M, \hat{\theta}, e]p_{t-1}[\omega_M]}{p[\text{outcome}|\omega_L, \hat{\theta}, e]p_{t-1}[\omega_L]}
\end{equation}

>>>>

# Framework 1

@@

$$\frac{p[h^t|\tilde{\theta}]}{p[h^t|\hat{\theta}]} > \alpha\geq 1$$

Where $h^t$ is the history of outcomes up to time $t$ and $\alpha$ is the switching threshold. By keeping track of the 
likelihood ratio, the agent can perform a \emph{Bayesian hypothesis test} and adopt the Dogmatic belief that best fits 
the data.\footnote{In a related problem @schwarstein proposes a similar updating procedure wich relies on the Bayesian 
hypothesis test and results in similar updating outcomes.}
Notice that if $\alpha \to \infty$, the behavior of the switcher will be indistinguishable from that of the Dogmatic modeler. 
In this sense, the switcher is a generalization of the dogmatic type.

By allowing the agent to keep track of the likelihoods and switching to an alternative type, the switcher can avoid the 
self-confirming equilibrium. However, if the prior belief on $omega$ is sufficiently tight around
a self-defeating equilibrium, the switcher might look identical to the dogmatic even in a case where $\alpha$ is not too large.
This happens because under the agent's prior, the likelihood ratio is unlikely to grow as fast as it is needed to escape 
the self-defeating equilibrium. In this cases we say that the misspecified belief is persistent. Hence, in order to be able 
to determine which of these two models provides a better explanation of the observed behaviors, we must focus on cases 
in which the prior is diffused enough and the sequence of realized outcomes is such that
the self-defeating equilibrium would not persist.

# Framework 2

@@

Bayesian agents always choose the effort level that maximizes their flow payoff by taking expectations over
their prior beliefs about $\theta$ and $\omega$. Since agents are myopic, even though all the parameters 
could be identified with enough variation in choices, a fully Bayesian agent might not learn their true type. This happens because they do not internalize the tradeoff between flow payoff
and learning and thus might not experiment enough to learn their type. An alternative to this approach is given 
by @Hestermann2021 and is discussed with the results.

@@

# A Unifying Example
The agent can be of one of 3 types: $\theta \in \{\theta_L, \theta_M, \theta_H\}$  with $\theta_H > \theta_M > \theta_L$.
They face an unknown exogenous success rate
$\omega \in \{\omega_L, \omega_M, \omega_H\}$ with $\omega_H>\omega_M>\omega_L$. Each of the values of $\omega$ is realized with 
equal probability. 
The agent knows the distribution of $\omega$ but not its realized value.

Denote the true type by $\theta^*$ and the true state by $\omega^*$. The agent holds some prior belief about $\theta$ 
\footnote{which is potentially misspecified as in the dogmatic and switcher cases discussed above} 
and chooses a binary gamble $e in \{e_L, e_M, e_H\}$. The agent observes whether the gamble is a success or a failure and 
gets a payoff of $1$ and if it is a success; they get $0$ otherwise.

The probability of success is increasing in both $\theta$ and $\omega$ and is fully described by the following table:

@@

Conditional on a type, the agent's flow payoff is maximized by choosing the gamble that 
matches the state. For example, if the value of $\omega$ is $\omega_H$, the agent's flow payoff is maximized by 
choosing $e_H$ and if the state is $\omega_L$ the flow payoff is maximized by choosing gamble $e_L$, regardless of the value of $\theta$. 
The agent myopically chooses gambles every period to maximize the flow payoff for $T<\infty$ periods.

After observing the outcome of each gamble, the agent updates their beliefs using some procedure and moves on to the next period.

Notice that both $\theta$ and $\omega$ can be identified from the outcomes if enough variation in the effort choices exists.
This can be seen by confirming that there is no pair of $\theta$ and $\omega$ such that the probability of success is the same for all effort 
choices. Thus, by changing the effort choice, the agent can learn both their type and the state if they observe enough outcomes.

In this example, for an agent with a dogmatic belief about their type, a self-defeating equilibrium is one in which the agent 
chooses an effort level that, under the true $\theta$, yields a frequency of success that is consistent with the agent's 
misspecified belief. That is $P[\text{sucess}|\theta^*, e^*] = P[\text{sucess}|\hat{\theta}, e^*]$ where $e^*$ is the agent's 
myopic optimal choice.

In the data-generating process described above, there are five such equilibria. For example, if the agent is of type $\theta_M$  but
mistakenly believes that he is of type $\hat{\theta}=\theta_H$ and the 
and $\omega^* = \omega_M$, when the effort chosen is $e_L$, the agent will observe a success with 45% chance. Because the agent 
dogmatically believes that their type is high, they will erroneously conclude that the rate is $\omega_L$. Under this belief,
the optimal action is $e_L$ which will continue to generate successes with $45%$ probability, further reinforcing
the incorrect belief. By doing so, the agent forgoes the payoff from gamble $e_M$ which would yield a success with $65\%$ chance.

By including self-confirming equilibria, the example captures the forces from each of the updating mechanisms discussed in the 
previous section and allows for direct comparison of all the theories. For realizations of $(\theta, \omega)$ for which there are 
self-confirming equilibria, the dogmatic agent will fall into the 
trap whereas the switcher will be able to escape it. Similarly, an agent with self-attribution bias will update their beliefs differently from an unbiased Bayesian, leading
them to choose different gambles. I exploit such cases in order to test which model is a better fit for how subjects behave in a 
laboratory experiment. \footnote{because the setting does not match that of @Heidhues2018, there will be situationf for which the theory 
does not provie a prediction. If such cases arise in the lab, they will not be used for the analysis.}

For this particular example,
